---
title: "Tooth Growth Analysis"
author: "cafkhami"
date: "Thursday, April 09, 2015"
output: html_document
---

##Exploratory Analysis
Lets graph a histogram of the tooth growth and separate out the colour by dose. 
```{r}
library(datasets)
library(ggplot2)
data(ToothGrowth)

g <- ggplot(ToothGrowth, aes(x = len, colour = as.factor(dose)))+geom_histogram(alpha = .20, binwidth=.5)
g# + facet_grid(. ~ dose)

```

This plot makes it look like there may be a difference in tooth growth based on the dose alone.

Now lets look at the effect of supplement type
```{r}
f <- ggplot(ToothGrowth, aes(x = len, colour = supp))+geom_histogram(alpha = .20, binwidth=.5)
f + facet_grid(. ~ supp)

```

This plot indicates that there may be a difference due to supplement type.  Let's explore further.

##Tests
Does dose really have an effect on the length of the tooth?  Lets determine the power of the assumption that the difference between a dose of 0.5 and 2.0 is greater than 5 

mu_2.0 - mu_0.5 > 5

Let's set up the data and preform a t.test.
```{r}
len_5 <- ToothGrowth$len[ToothGrowth$dose == 0.5]
len_2 <- ToothGrowth$len[ToothGrowth$dose == 2]
delta <- 5
n = length(len_5)
sd = sd(len_2-len_5)

power.t.test(n,delta,sd,type = "paired", alt = "one.sided")$power

t.test(ToothGrowth$supp, ToothGrowth$len, paired = FALSE)$conf
```